---
title: ""
output: 
  html_document: 
    css: boosh.css
---

### False discovery and friends in psycholinguistics 

**[Timothy Leffel](http://lefft.xyz), NORC at U Chicago**   
[*draft, updated july23/2017*]
<hr style="font-size: 20px; color: #0f422f; font-weight: bold; border-bottom: 2px solid DimGray;">


### 

**Abstract.** 
Simulating RT-based psycholinguistics experiments to understand the role of analysis decisions and power in type-i/-ii errors, false discovery rate, and related problems. 


#### set up environment

<!-- import google fonts -->
<link rel="stylesheet" type="text/css"
  href="https://fonts.googleapis.com/css?family=Open+Sans:300,400,400i,700">
<link href="https://fonts.googleapis.com/css?family=Roboto+Mono:300,400" rel="stylesheet">


```{r warning=FALSE, message=FALSE}
# NOTE: list of TODO items at bottom of script

# set global chunk options
knitr::opts_chunk$set(
  fig.width=10, fig.height=7, warning=FALSE, message=FALSE
)

# load dependencies
library("dplyr"); library("magrittr"); library("ggplot2"); library("gridExtra")

# set plot theme
theme_set(theme_minimal(14))

# set random seed for reproducibility
set.seed(1234)

# load simulation functions etc.
source("functions.r")
```


#### simulate a simple experiment

Suppose we're doing an RT experiment + we have just related/unrelated conditions. Further suppose everything is matched on freq/length/etc. (see `functions.r` for definition of `simulate_experiment()` -- will make design more complex/realistic in next iteration).

```{r simulate_expt}
# simulate the experiment a single time
dat <- simulate_experiment()
```

The data look like this: 

```{r}
# display 10 randomly sampled rows of the data 
knitr::kable(dat[sample(seq(len=nrow(dat)), size=10), ], digits=2)
```

We've already specified the number of outlier subjs (in the function def'n), but we'll want to have a couple params around for stuff below. This needs to be cleaned up. 

**TODO (important)**: fix this for clarity -- *need to set global params for the subsetting etc. below. These are already fed into `simulate_experiment()` as default args, but should be passed explicitly from the global environment for transparency. But need to improve/transparent-ize the argument passing structure in the sim pipeline before that is possible).*

```{r}
# set params for subsetting below [HACK -- FIX THIS SOON]
n_outlier    <- length(unique(dat$subj[dat$is_outlier=="_outlier"]))
outlier_subj <- unique(dat$subj[dat$is_outlier=="_outlier"])
```


#### plot the raw RT's

```{r plot_rez}
# plot the resutls by condition and outlier status 
# (options: rt, rt_log, rt_ctd, rt_zsc)
plot_results(dat, rt_col="rt")
```

*(and summarize them -- need to generalize summary func before can do summaries for other RT measures)*

**TODO:** generalize summary func (prob need to implement it w `enquo()` for smoothness ack)

```{r summarize_rez}
# summarize the results by relevant grouping vars
tabs <- summarize_results(dat)

tabs$summary_by_cond
tabs$summary_by_outlier
tabs$summary_by_outlier_cond
```


#### plot the raw RT's, with outliers excluded

```{r}
# plot the raw rt's, but with outlier subj's excluded
plot_results(dat[dat$is_outlier=="reggie", ], rt_col="rt")

# summarize them too
summarize_results(dat[dat$is_outlier=="reggie", ])$summary_by_cond
```


#### fit some models [STILL NEEDS WRITTEN]

```{r analysis_chunk} 
# [THIS IS THE CRUCIAL PART RITE HERE]
```

**TODO:** encapsulate code in next chunk into a func
```{r results="hold"}
# establish + quantify the effect of condition
fit <- lm(rt ~ cond, data=dat)
knitr::kable(summary(fit)$coefficients, caption="<br>model: RT ~ condition")

# add outlier:cond interaction as a predictor + fit
fit_outlier <- lm(rt ~ cond + cond:is_outlier, data=dat)
knitr::kable(summary(fit_outlier)$coefficients, 
             caption="<br><br>model: RT ~ condition + condition:outlier")

# evaluate relative fit between the reggie mod + one w outlier interaction term
anova(fit, fit_outlier) # [c("F","Pr(>F)")]
```


```{r}
# also: exclude outliers to see how effect size changes relative to `fit`
fit_excl <- lm(rt ~ cond, data=dat[!dat$subj %in% outlier_subj, ])
knitr::kable(summary(fit_excl)$coefficients)
```

<hr><hr>
<br><br><br><br><br><br>



#### notes to self 

```{r}
# can use either strategy for passing metadata
# comment(dat) <- "boosh!"
# attr(dat, "tim_note") = "boosh!"

# NOTE: also fit the model on logged rt's 
# (can use scaled to get sd units too)
# fit_log <- lm(rt_log ~ cond, data=dat); summary(fit_log)

# flow of data in sim setup is: 
# generate_data() %>% introduce_outliers() %>% transform_rts()

# NOTE: the attributes of dat$rt_zsc are mean and sd ("center" and "scale")

# width of individual kable()s, use arg: table.attr = "style='width:XXX%;'"
```

<!-- don't want to evaluate this chunk -- might be useful once we start modeling real data tho -->
```{r plot_scaled_rts, eval=FALSE, include=FALSE, echo=FALSE}
#### code scraps, misc etc.

# don't want to evaluate this chunk -- 
# might be useful once we start modeling real data tho

#### plot the logged RT's

# plot the logged rt's
plot_results(dat, rt_col="rt_log")

#### plot the centered RT's

# plot the centered rt's
plot_results(dat, rt_col="rt_ctd")

#### plot the z-scored RT's

# plot the logged rt's
plot_results(dat, rt_col="rt_zsc")
```



<br><br><br>
<hr style="size: 40px; color: orange;">




